### 2018-01-15 submission/manuscript.Rmd
add completed intro with citations

Over recent years there has been an increasing focus on standardizing methodological approaches in microbiota research []. In particular, a lot of the attention of this standardization push has been invested in investigating ways that 16S rRNA gene surveys can be made more replicatable []. Although 16S rRNA gene sequencing has been much maligned for introduced bias, many of these same considerations also affect metagenomic sequencing [@metagenomcis_bias_Nayfach_2016]. Between the two approaches similar bias considerations include, but are not limited to, preservation media, storage conditions, DNA extraction kit, PCR, and sequece library preparation. Thus what is identified as a source of bias in 16S rRNA gene sequencing for these overlapping considerations will also influence metagenomic sequencing results.

The most commonly studied biases are by far the skews introduced to the data via preservation media used, storage conditions, and DNA extraction kits chosen. The study of these specific biases has become so large, aggregating them all together has become a difficult task and some researchers provide resources to actively track new findings (e.g. Microbiome Digest - https://microbiomedigest.com/microbiome-papers-collection/microbiome-techniques/sample-storage/). Within the literature DNA extraction kits and storage condition chosen have consistently been shown to add bias to downstream analysis []. However, the current literature on preservation media has been more mixed, with some studies showing biases while others do not []. Although these are important sources of bias they are not the only sources that should be critically examined. The type of DNA polymerase chosen could have a wide ranging affect on downstream results due to error rates and chimeras that may not be easily resolved using bioinformatic approaches.

>>>>

Over recent years there has been an increasing focus on standardizing methodological approaches in microbiota research [@review_Kim_2017; @review_Hugerth_2017]. In particular, a lot of the attention of this standardization push has been invested in investigating ways that 16S rRNA gene surveys can be made more replicatable [@contamination_Salter2014; @taq_Gohl_2016; @storage_Lauber_2010; @preservation_Song_2016]. Although 16S rRNA gene sequencing has been much maligned for introduced bias, many of these same considerations also affect metagenomic sequencing [@metagenomcis_bias_Nayfach_2016; @metagenomcis_bias_Costea_2017].  Between the two approaches similar bias considerations include, but are not limited to, preservation media, storage conditions, DNA extraction kit, PCR, and sequece library preparation. Thus what is identified as a source of bias in 16S rRNA gene sequencing for these overlapping considerations will also influence metagenomic sequencing results.

The most commonly studied biases are by far the skews introduced to the data via preservation media used, storage conditions, and DNA extraction kits chosen. The study of these specific biases has become so large, aggregating them all together has become a difficult task and some researchers provide resources to actively track new findings (e.g. Microbiome Digest - https://microbiomedigest.com/microbiome-papers-collection/microbiome-techniques/sample-storage/). Within the literature DNA extraction kits have consistently been shown to add bias to downstream analysis [@contamination_Salter2014; @metagenomcis_bias_Costea_2017]. However, the current literature on preservation media and storage conditions has been more mixed, with some studies showing biases while others do not [@preservation_Song_2016; @preservation_Luo_2016; @preservation_Dominianni_2014; @storage_Lauber_2010; @storage_Bassis_2017; @preservation_Sinha_2015]. Although these are important sources of bias they are not the only sources that should be critically examined. The type of DNA polymerase chosen could have a wide ranging affect on downstream results due to error rates and chimeras that may not be easily resolved using bioinformatic approaches.

### 2018-01-15 submission/manuscript.tex
add completed intro with citations

Over recent years there has been an increasing focus on standardizing
methodological approaches in microbiota research {[}{]}. In particular,
a lot of the attention of this standardization push has been invested in
investigating ways that 16S rRNA gene surveys can be made more
replicatable {[}{]}. Although 16S rRNA gene sequencing has been much
maligned for introduced bias, many of these same considerations also
affect metagenomic sequencing (Nayfach \& Pollard, 2016). Between the
two approaches similar bias considerations include, but are not limited
to, preservation media, storage conditions, DNA extraction kit, PCR, and
sequece library preparation. Thus what is identified as a source of bias
in 16S rRNA gene sequencing for these overlapping considerations will
also influence metagenomic sequencing results.

The most commonly studied biases are by far the skews introduced to the
data via preservation media used, storage conditions, and DNA extraction
kits chosen. The study of these specific biases has become so large,
aggregating them all together has become a difficult task and some
researchers provide resources to actively track new findings
(e.g.~Microbiome Digest -
\url{https://microbiomedigest.com/microbiome-papers-collection/microbiome-techniques/sample-storage/}).
Within the literature DNA extraction kits and storage condition chosen
have consistently been shown to add bias to downstream analysis {[}{]}.
However, the current literature on preservation media has been more
mixed, with some studies showing biases while others do not {[}{]}.
Although these are important sources of bias they are not the only
sources that should be critically examined. The type of DNA polymerase
chosen could have a wide ranging affect on downstream results due to
error rates and chimeras that may not be easily resolved using
bioinformatic approaches.

@@

\hypertarget{refs}{}
\hypertarget{ref-benjamini_controlling_1995}{}
Benjamini Y., Hochberg Y. 1995. Controlling the false discovery rate: A
practical and powerful approach to multiple testing. \emph{Journal of
the Royal Statistical Society. Series B (Methodological)} 57:289--300.

@@

@@

@@

@@

@@

>>>>

Over recent years there has been an increasing focus on standardizing
methodological approaches in microbiota research (Kim et al., 2017;
Hugerth \& Andersson, 2017). In particular, a lot of the attention of
this standardization push has been invested in investigating ways that
16S rRNA gene surveys can be made more replicatable (Lauber et al.,
2010; Salter et al., 2014; Song et al., 2016; Gohl et al., 2016).
Although 16S rRNA gene sequencing has been much maligned for introduced
bias, many of these same considerations also affect metagenomic
sequencing (Nayfach \& Pollard, 2016; Costea et al., 2017). Between the
two approaches similar bias considerations include, but are not limited
to, preservation media, storage conditions, DNA extraction kit, PCR, and
sequece library preparation. Thus what is identified as a source of bias
in 16S rRNA gene sequencing for these overlapping considerations will
also influence metagenomic sequencing results.

The most commonly studied biases are by far the skews introduced to the
data via preservation media used, storage conditions, and DNA extraction
kits chosen. The study of these specific biases has become so large,
aggregating them all together has become a difficult task and some
researchers provide resources to actively track new findings
(e.g.~Microbiome Digest -
\url{https://microbiomedigest.com/microbiome-papers-collection/microbiome-techniques/sample-storage/}).
Within the literature DNA extraction kits have consistently been shown
to add bias to downstream analysis (Salter et al., 2014; Costea et al.,
2017). However, the current literature on preservation media and storage
conditions has been more mixed, with some studies showing biases while
others do not (Lauber et al., 2010; Dominianni et al., 2014; Sinha et
al., 2015; Song et al., 2016; Luo et al., 2016; Bassis et al., 2017).
Although these are important sources of bias they are not the only
sources that should be critically examined. The type of DNA polymerase
chosen could have a wide ranging affect on downstream results due to
error rates and chimeras that may not be easily resolved using
bioinformatic approaches.

@@

\hypertarget{refs}{}
\hypertarget{ref-storage_Bassis_2017}{}
Bassis CM., Nicholas M. Moore., Lolans K., Seekatz AM., Weinstein RA.,
Young VB., Hayden MK. 2017. Comparison of stool versus rectal swab
samples and storage conditions on bacterial community profiles.
\emph{BMC Microbiology} 17. DOI:
\href{https://doi.org/10.1186/s12866-017-0983-9}{10.1186/s12866-017-0983-9}.

\hypertarget{ref-benjamini_controlling_1995}{}
Benjamini Y., Hochberg Y. 1995. Controlling the false discovery rate: A
practical and powerful approach to multiple testing. \emph{Journal of
the Royal Statistical Society. Series B (Methodological)} 57:289--300.

@@

\hypertarget{ref-metagenomcis_bias_Costea_2017}{}
Costea PI., Zeller G., Sunagawa S., Pelletier E., Alberti A., Levenez
F., Tramontano M., Driessen M., Hercog R., Jung F-E., Kultima JR.,
Hayward MR., Coelho LP., Allen-Vercoe E., Bertrand L., Blaut M., Brown
JRM., Carton T., Cools-Portier S., Daigneault M., Derrien M., Druesne
A., Vos WM de., Finlay BB., Flint HJ., Guarner F., Hattori M., Heilig
H., Luna RA., Hylckama Vlieg J van., Junick J., Klymiuk I., Langella P.,
Chatelier EL., Mai V., Manichanh C., Martin JC., Mery C., Morita H.,
O'Toole PW., Orvain C., Patil KR., Penders J., Persson S., Pons N.,
Popova M., Salonen A., Saulnier D., Scott KP., Singh B., Slezak K.,
Veiga P., Versalovic J., Zhao L., Zoetendal EG., Ehrlich SD., Dore J.,
Bork P. 2017. Towards standards for human fecal sample processing in
metagenomic studies. \emph{Nature Biotechnology}. DOI:
\href{https://doi.org/10.1038/nbt.3960}{10.1038/nbt.3960}.

\hypertarget{ref-preservation_Dominianni_2014}{}
Dominianni C., Wu J., Hayes RB., Ahn J. 2014. Comparison of methods for
fecal microbiome biospecimen collection. \emph{BMC Microbiology} 14:103.
DOI:
\href{https://doi.org/10.1186/1471-2180-14-103}{10.1186/1471-2180-14-103}.

@@

\hypertarget{ref-review_Hugerth_2017}{}
Hugerth LW., Andersson AF. 2017. Analysing microbial community
composition through amplicon sequencing: From sampling to hypothesis
testing. \emph{Frontiers in Microbiology} 8. DOI:
\href{https://doi.org/10.3389/fmicb.2017.01561}{10.3389/fmicb.2017.01561}.

@@

\hypertarget{ref-review_Kim_2017}{}
Kim D., Hofstaedter CE., Zhao C., Mattei L., Tanes C., Clarke E., Lauder
A., Sherrill-Mix S., Chehoud C., Kelsen J., Conrad M., Collman RG.,
Baldassano R., Bushman FD., Bittinger K. 2017. Optimizing methods and
dodging pitfalls in microbiome research. \emph{Microbiome} 5. DOI:
\href{https://doi.org/10.1186/s40168-017-0267-5}{10.1186/s40168-017-0267-5}.

@@

\hypertarget{ref-storage_Lauber_2010}{}
Lauber CL., Zhou N., Gordon JI., Knight R., Fierer N. 2010. Effect of
storage conditions on the assessment of bacterial community structure in
soil and human-associated samples. \emph{FEMS Microbiology Letters}
307:80--86. DOI:
\href{https://doi.org/10.1111/j.1574-6968.2010.01965.x}{10.1111/j.1574-6968.2010.01965.x}.

\hypertarget{ref-preservation_Luo_2016}{}
Luo T., Srinivasan U., Ramadugu K., Shedden KA., Neiswanger K., Trumble
E., Li JJ., McNeil DW., Crout RJ., Weyant RJ., Marazita ML., Foxman B.
2016. Effects of specimen collection methodologies and storage
conditions on the short-term stability of oral microbiome taxonomy.
\emph{Applied and Environmental Microbiology} 82:5519--5529. DOI:
\href{https://doi.org/10.1128/aem.01132-16}{10.1128/aem.01132-16}.

@@

\hypertarget{ref-preservation_Sinha_2015}{}
Sinha R., Chen J., Amir A., Vogtmann E., Shi J., Inman KS., Flores R.,
Sampson J., Knight R., Chia N. 2015. Collecting fecal samples for
microbiome analyses in epidemiology studies. \emph{Cancer Epidemiology
Biomarkers \& Prevention} 25:407--416. DOI:
\href{https://doi.org/10.1158/1055-9965.epi-15-0951}{10.1158/1055-9965.epi-15-0951}.

\hypertarget{ref-preservation_Song_2016}{}
Song SJ., Amir A., Metcalf JL., Amato KR., Xu ZZ., Humphrey G., Knight
R. 2016. Preservation methods differ in fecal microbiome stability,
affecting suitability for field studies. \emph{mSystems} 1:e00021--16.
DOI:
\href{https://doi.org/10.1128/msystems.00021-16}{10.1128/msystems.00021-16}.
